### 2010-10-11 Thesis/Chapter7.tex
- Update 11 October -
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@184 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\cite{Volino06} suggested an alternative somewhere between mass-spring and geometrically derived approaches. They compute a bending vector that represents the bending of the surface through  a simple linear combination of particule positions and they distribute this vector as particule forces according to the bending stiffness. The authors claimed much better accuracy than mass-spring models (particulary when dealing with low curvature and high bending stiffness) with similar computation time (see \fig{chap7:fig-dress} for an illustration). 
%
\begin{figure}[ht]
\begin{center}
\includegraphics[width=10cm]{chapter7/dress.pdf}
\caption[Simulation of complete garments that have large bending stiffness]{The model of \citeauthor{Volino06} allows the simulation of complete garments with material that may have fairly large bending stiffness. Image courtesy of \cite{Volino06}.}
\label{chap7:fig-dress}
\end{center}
\end{figure}

@@

Shell finite element modelling followed the same trend than solid objects and co-rotational formulations were introduced to simplify strain expressions while allowing large rotations (but small strain) (see page \pageref{chap4:corotationalMethods} for more details). For instance, \cite{Etzmuss03} used the linear Cauchy strain tensor but applied a polar decomposition on the deformation gradient to allow large deformations. The bending energy is separated from the membrane formulation and most of material parameters were measured from experiments on actual pieces of clothing. As an example, the computation of a shirt consisting of $ 8\,300 $ triangles took an average of less than $ 3\,$s per animation frame with a time step of $ 0.02\,$s (see an illustration \fig{chap7:fig-shirt}). While still being far from achieving real-time, the intention of the authors was clearly to propose an efficiency model for cloth simulation. 
%
\begin{figure}[ht]
\begin{center}
\includegraphics[width=12cm]{chapter7/shirt.pdf}
\caption[A shirt shown with different material properties]{A shirt shown with different material properties: (a) wool/viscose, (b) wool and (c) polyester/polyacrylics/acetate. Image courtesy of \cite{Etzmuss03}.}
\label{chap7:fig-shirt}
\end{center}
\end{figure}

>>>>

\cite{Volino06} suggested an alternative somewhere between mass-spring and geometrically derived approaches. They compute a bending vector that represents the bending of the surface through  a simple linear combination of particule positions and they distribute this vector as particule forces according to the bending stiffness. The authors claimed much better accuracy than mass-spring models (particulary when dealing with low curvature and high bending stiffness) with similar computation time (see \fig{chap7:fig-dress} for an illustration). 
%
\begin{figure}[ht]
\begin{center}
\includegraphics[width=10cm]{chapter7/dress.pdf}
\caption[Simulation of complete garments that have large bending stiffness]{The model of Volino and Magnenat-Thalmann allows the simulation of complete garments with material that may have fairly large bending stiffness. Image courtesy of \cite{Volino06}.}
\label{chap7:fig-dress}
\end{center}
\end{figure}

@@

Shell finite element modelling followed the same trend than solid objects and co-rotational formulations were introduced to simplify strain expressions while allowing large rotations (but small strain, see page \pageref{chap4:corotationalMethods} for more details). For instance, \cite{Etzmuss03} used the linear Cauchy strain tensor but applied a polar decomposition on the deformation gradient to allow large deformations. The bending energy is separated from the membrane formulation and most of material parameters were measured from experiments on actual pieces of clothing. As an example, the computation of a shirt consisting of $ 8\,300 $ triangles took an average of less than $ 3\,$s per animation frame with a time step of $ 0.02\,$s (see an illustration \fig{chap7:fig-shirt}). While still being far from achieving real-time, the intention of the authors was clearly to propose an efficiency model for cloth simulation. 
%
\begin{figure}[ht]
\begin{center}
\includegraphics[width=12cm]{chapter7/shirt.pdf}
\caption[A shirt shown with different material properties]{A shirt shown with different material properties: (a) wool/viscose, (b) wool and (c) polyester/polyacrylics/acetate. Image courtesy of \cite{Etzmuss03}.}
\label{chap7:fig-shirt}
\end{center}
\end{figure}

### 2010-10-11 Thesis/Chapter9.tex
- Update 11 October -
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@184 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

From there, our approach to measure the rotations around local x and y axes is simple. At runtime and for each vertex, we compute the rotation that we need to transform the z axis of the triangle's frame into the z axis of the vertex's frame (see listing \ref{chap9:qDiffZ}). The differences with the same quantities computed for the rest position during the initialisation lead to the desired displacements.

\begin{lstlisting}[caption={Method used to compute the rotation of the z axis between the frames embodied by two quaternions},label=chap9:qDiffZ]
inline Quat qDiffZ(const Quat& Qvertex, const Quat& Qtriangle)
{
    // dQ is the quaternion that embodies the rotation between the z axis of the vertex and the z axis of the local triangle's frame (in local space)
    Quat dQ;

@@

    // v = z axis of the vertex's frame is expressed into world space
    Vec3d v = Qvertex.rotate(Vec3d(0.0, 0.0, 1.0));

@@

    // Axis of rotation between the 2 vectors u and v lies into the plan of the 2 vectors
    Vec3d axis = cross(u, v0);

@@

It is worth noting that a high resolution mesh may be obtained by recursively subdividing each triangle. However, in this case the vertices are closely related to the structure of the mechanical mesh and rendering artifacts may appear. As we can notice on \fig{chap9:fig-rendering}, the quality of the visualisation may be improved by using a totally different mesh.
%
%\begin{figure}
%	\begin{center}
%		\begin{tabular}{cc}
%			\includegraphics[width=6cm]{chapter9/sub_grid_red.png} & \includegraphics[width=6cm]{chapter9/irregular_grid_red.png} \\
%			\includegraphics[width=6cm]{chapter9/sub_grid_mesh.png} & \includegraphics[width=6cm]{chapter9/irregular_grid_mesh.png} \\
%			Subdivided triangles & Irregular grid			
%		\end{tabular}
%	\end{center}
%	\label{chap9:fig-rendering}
%	\caption{We can notice some rendering artifacts when we use subdivided triangles. Using an irregular mesh instead reduces the impact.}
%\end{figure}
%
\begin{figure}[ht]
\centering 
\subfloat[Subdivided triangles]{\includegraphics[width=5cm]{chapter9/sub_grid_red.png}}
\hspace{1cm}
\subfloat[Irregular grid]{\includegraphics[width=5cm]{chapter9/irregular_grid_red.png}} \\
\subfloat[Subdivided triangles]{\includegraphics[width=5cm]{chapter9/sub_grid_mesh.png}}
\hspace{1cm}
\subfloat[Irregular grid]{\includegraphics[width=5cm]{chapter9/irregular_grid_mesh.png}}
\caption[Visual artifacts using triangle subdivisions]{In some situations, some rendering artifacts may appear when we use subdivided triangles to render the shells (a,c). Using an irregular mesh which is not closely related to the structure of the mechanical mesh reduces the impact (b,d).}
\label{chap9:fig-rendering}
\end{figure}

>>>>

From there, our approach to measure the rotations around local x and y axes is simple. At runtime and for each vertex, we compute the rotation that we need to transform the z axis of the triangle's frame into the z axis of the vertex's frame (see listing \ref{chap9:qDiffZ}). The differences with the same quantities computed for the rest position during the initialisation lead to the desired displacements. 
%
\begin{lstlisting}[caption={[Method to compute the rotation between two quaternions]Method used to compute the rotation of the z axis between the frames embodied by two quaternions},label={chap9:qDiffZ},frame=shadowbox,rulesepcolor=\color{black}]
inline Quat qDiffZ(const Quat& Qvertex, const Quat& Qtriangle)
{
    // dQ is the quaternion that embodies the rotation between 
    // the z axis of the vertex and the z axis of the local 
    // triangle's frame (in local space)
    Quat dQ;

@@

    // v = z axis of the vertex's frame is expressed into 
    // world space
    Vec3d v = Qvertex.rotate(Vec3d(0.0, 0.0, 1.0));

@@

    // Axis of rotation between the 2 vectors u and v lies 
    // into the plan of the 2 vectors
    Vec3d axis = cross(u, v0);

@@

It is worth noting that a high resolution mesh may be obtained by recursively subdividing each triangle. However, in this case the vertices are closely related to the structure of the mechanical mesh and rendering artifacts may appear. As we can notice on \fig{chap9:fig-rendering}, the quality of the visualisation may be improved by using a totally different mesh.
%
\begin{figure}[ht]
\centering 
\subfloat[Subdivided triangles]{\includegraphics[width=5.5cm]{chapter9/sub_grid_red.png}}
\hspace{1cm}
\subfloat[Irregular grid]{\includegraphics[width=5.5cm]{chapter9/irregular_grid_red.png}} \\
\subfloat[Subdivided triangles]{\includegraphics[width=5.5cm]{chapter9/sub_grid_mesh.png}}
\hspace{1cm}
\subfloat[Irregular grid]{\includegraphics[width=5.5cm]{chapter9/irregular_grid_mesh.png}}
\caption[Visual artifacts using triangle subdivisions]{In some situations, some rendering artifacts may appear when we use subdivided triangles to render the shells (a,c). Using an irregular mesh which is not closely related to the structure of the mechanical mesh reduces the impact (b,d).}
\label{chap9:fig-rendering}
\end{figure}
